---
title: "Pokemon API"
author: "Ryan Bunn"
---

This page is a vignette showing the retrevial of data from the Pokemon API found at [https://pokeapi.co/]

#Required Packages
For obtaining information from the Pokemon API and the subsequent data analysis the following packages were used

`tidyverse`
`httr`
`jsonlite`
`ggplot2`

```{r libraries, echo=FALSE}
library(tidyverse)
library(httr)
library(jsonlite)
library(ggplot2)
```

#API Functions
##Results Dataframe
As the API functions are designed to be used on a single variable input these functions can take in the results from a `lapply` of the API accessing functions and turn them into a nice dataframe.
```{r dataframe}
  ResultsToDataframe<- function(object){
    object<-as.numeric(as.character(object))
    object<-unlist(object)
    object<-as.data.frame(object)
    object[sapply(object, is.na)] = 0
    return(object)
  }
  depth2<- function(value){
    value <-unlist(value)
    value <-value[c(TRUE,FALSE)]
    return(value)
  }
  removeNull <- function(x, func=function(x) if(is.null(x)) NA else x)
{
  if(is.list(x))
  {
    lapply(x, removeNUll, func)
  } else
  {
    func(x)
  }
}
```

##Berries
This function is designed to extract information on berries, the `type` variable allows for information to be extracted based upon berry type, firmness, or flavor. In addition the `specific` variable when not specificed extracts data on the different subvariables. When `specific` is specified, as one of the subvariables, more detailed information can be extracted. In addition, the `val` variable allows for highly detailed information to be extracted about a particular berry. Finally the limit variable allows us to limit the number of results as needed.
```{r berry}
berry <- function(type,specific="",val="",limit=100){
  limit = as.character(limit)
  if(!is_character(type)){
    stop("Error: Must be string")
  }
  base <- "https://pokeapi.co/api/v2"
  if(type=="berry"){
    hyperlink <- paste0(base,"/berry/",specific,"/?limit=",limit)
    a<-GET(hyperlink)
    data <- fromJSON(rawToChar(a$content))
    if(val!=""){return(data[[val]])}
    if(specific != ""){
    return(data)}
    else{
      return(data$results)}
  }
    else if (type=="firmness"){
      hyperlink <- paste0(base,"/berry-firmness/",specific,"/?limit=",limit)
      a<-GET(hyperlink)
      data <- fromJSON(rawToChar(a$content))
      if(specific != ""){
    return(data$berries)}
    else{
      return(data$results)}
  }
      else if(type=="flavor"){
        hyperlink <- paste0(base,"/berry-flavor/",specific,"/?limit=",limit)
        a<-GET(hyperlink)
        data <- fromJSON(rawToChar(a$content))
        if(specific != ""){
    return(data$berries)}
    else{
      return(data$results)}
  }
        else{
          stop("Error: Must be one of berry,firmness,flavor")
        }
}
```



##Moves
This function is designed to extract information on moves, the `type` variable allows for information to be extracted based upon move, category, or class. In addition the `specific` variable when not specificed extracts data on the different subvariables. When `specific` is specified, as one of the subvariables, more detailed information can be extracted. The `val` variable allows for highly detailed information to be extracted about a particular move. Finally the limit variable allows us to limit the number of results as needed.
```{r move}
move <- function(type,specific="",val="",val2="",limit=1000){
  if(!is_character(type) || !is_character(val)){
    stop("Error: Must be string")
  }
  if(type == "move"){
  base <- "https://pokeapi.co/api/v2/move/"
  link<-paste0(base,specific,"/?limit=",limit)
  a<- GET(link)
  r<- fromJSON((rawToChar(a$content)))
  if(val != ""){return(unlist(r[[val]]))}
  if(specific != ""){return(r)}
  return(r$results)
  }else if(type == "category"){
      base <- "https://pokeapi.co/api/v2/move-category/"
      link<-paste0(base,specific,"/?limit=",limit)
      a<- GET(link)
      r<- fromJSON((rawToChar(a$content)))
      if(specific != "") return(r$moves)
      else{return(r$results)}
  } else  if(type == "class"){
    base <- "https://pokeapi.co/api/v2/move-damage-class/"
    link<-paste0(base,specific,"/?limit=",limit)
    a<- GET(link)
    r<- fromJSON((rawToChar(a$content)))
    if(specific != ""){
      if(val != ""){return(r[[val]])}
      return(r$moves)}
    else{return(r$results)}
  } else {
    stop("Error: Must be one of move,category,class")
  }
}
```


##Pokemon
This function is designed to extract information on pokemon, the `type` variable allows for information to be extracted based upon ability, pokemon, or stat of interest. In addition the `specific` variable when not specificed extracts data on the different subvariables. When `specific` is specified, as one of the subvariables, more detailed information can be extracted.The `val` variable allows for highly detailed information to be extracted about a particular pokemon. `val2` and `val3` are used with the pokemon endpoint to travel deeper into the provided data. In addition, the `inc` variable allows the user to specify if they would like the stats results to be in increasing order. Finally the limit variable allows us to limit the number of results as needed.
```{r pokemon}
  pokemon <- function(type,specific="",val="",val2="",val3="",inc=TRUE,limit=1000){
    if(!is_character(type)){
      stop("Error: Must be string")
    } 
  
    if(type=="ability"){
      base <- "https://pokeapi.co/api/v2/ability/"
      link <- paste0(base,specific,"/?limit=",limit)
      a <- GET(link)
      r<- fromJSON(rawToChar(a$content))
      if(specific != ""){return(r$pokemon)}
      return(r)
    }
    if(type == "pokemon"){
      base <- "https://pokeapi.co/api/v2/pokemon/"
      link <- paste0(base,specific,"/?limit=",limit)
      a <- GET(link)
      r<- fromJSON(rawToChar(a$content))
      if(val !=""){
        if(val2 != ""){
          if(val3 !=""){return(r[[val]][[val2]][[val3]])}
          
          return(r[[val]][val2])}
        return(r[[val]])}
      if(specific!= ""){return(r$weight)}
      else {return(r$results)}
    }
    if(type == "stats"){
      base <- "https://pokeapi.co/api/v2/stat/"
      link <- paste0(base,specific,"/?limit=",limit)
      a <- GET(link)
      r<- fromJSON(rawToChar(a$content))
      if(specific!=""){
        if(inc){return(r$affecting_moves$increase)}
        else{return(r$affecting_moves$decrease)}
      }
      else{return(r$results)}
    }
  }
  
```


##Wrapper
This wrapper function is designed so that only a single function needs to be called with the function of interest and any other inputs specified.
```{r wrapper}
  API_Wrapper <- function(fun,...){
    if(!is_character(fun)){
    stop("Error: Must be string for function")
  }
    if(fun=="berry"){
      berry(...)
    }
      else if(fun=="move"){
        move(...)
      }
        else if(fun=="pokemon"){
          pokemon(...)
        }
          else{
            stop("Error: fun must be one of berry, move, pokemon")}
  }
```

#Exploratory Data Analysis
##Moves
We are interested in specifics about the various moves in the pokemon game. We start by collecting a list of the first 250 moves and then determining the accuracy, power, pp, and damage class.
```{r moves}
all_moves<-API_Wrapper("move",type="move",limit=1000)[,1]
#all_moves
accuracy<-sapply(X=all_moves,API_Wrapper,fun="move",type="move",val="accuracy")
accuracy <- ResultsToDataframe(accuracy)

power <- sapply(X=all_moves,API_Wrapper,fun="move",type="move",val="power")
power <- ResultsToDataframe(power)

pp <- sapply(X=all_moves,API_Wrapper,fun="move",type="move",val="pp")
pp<-ResultsToDataframe(pp)

class <- sapply(X=all_moves, API_Wrapper,fun="move",type="move",val="damage_class")
class <- removeNull(class)
class <- sapply(class,`[`,"name")
class <- as.data.frame(class)

data<- cbind(accuracy,power,pp,class)
colnames(data)<- c("accuracy","power","pp","class")

```

Next we can create some a variable that gives us an insight into the effective power of a move by accounting for limitations in power, accuracy and/or number of available uses. It may be benificial to see how many of our responses fall into particular categories for class Finally we can also look at some numerical summarises of the variables based upon the class of the move.
```{r datawork}
data <- data %>% mutate(effectivePower = power*accuracy)

table(data$class)

data %>% group_by(class) %>% summarise(avg_power = mean(power), sd_power=sd(power),avg_accuracy=mean(accuracy),sd_accuracy=sd(accuracy),avg_pp=mean(pp),sd_pp=sd(pp))

```
Observing the output contingency table we see that there are more physical moves than special and status moves, it may be interesting to see how the moves further breakdown in future analysis. The numerical summaries show us an NA section for class which may mean the API is missing data or there is a undefined class of moves. It may be beneficial to investigate the reason for the NA later.  



We can then make some comparisons between the effective power and the number of times an ability can be used for each class by looking at a scatter plot
```{r scatter}
  g<-ggplot(data,aes(effectivePower,pp))
  g+geom_point(aes(color=class))+labs(x="Effective Power",y="PP",title = "Effective Power vs PP by Class")
```
Observing the scatterplot we see that a high effective power seems to guarantee a low PP value, but a low PP value does not guarentee a high effective power. We also see that the status class sticks to the 0 effective power, which mirrors the results seen summary table before.


Next we can look at the distribution of accuracy across moves 
```{r histogram}
  h<- ggplot(data,aes(accuracy))
  h+geom_histogram(fill="red")+labs(title="Histogram of Accuracy")
```
Our histogram tells us that the most moves generally fall under 0 accuracy(accuracy irrelevent) or 100 accuracy and very few elements have an accuracy in between. It may be interesting to see in future analysis how the type and/or effective power factor into accuracy values.


#Berries
Next we are interested in looking at the aspects that berrys in the pokemon games have like their size, growth speed, harvest quantity, firmness type, and flavor.
```{r berries}
berries <- API_Wrapper("berry",type="berry",limit=1000)[,1]
size <- sapply(X=berries, API_Wrapper,fun="berry",type="berry",val="size")
size<-ResultsToDataframe(size)

growth <- sapply(X=berries, API_Wrapper,fun="berry",type="berry",val="growth_time")
growth<-ResultsToDataframe(growth)

harvest <- sapply(X=berries, API_Wrapper,fun="berry",type="berry",val="max_harvest")
harvest<-ResultsToDataframe(harvest)

firm <- sapply(X=berries,API_Wrapper,fun="berry",type="berry",val="firmness")
firm<-depth2(firm)

gift <- sapply(X=berries,API_Wrapper,fun="berry",type="berry",val="natural_gift_type")
gift <- depth2(gift)

data2 <- cbind(size,growth,harvest,firm,gift)
colnames(data2) <- c("size","growth","harvest","firmness","natural_gift")

```

We can then create 2 new variables to put growth time into class and then a variable for the number harvested for growth time unit. Next we look at parrallel boxplots of the size based upon the new growth time class.
```{r boxplot}
data2 <- data2 %>% mutate(g_class=if_else(growth >15,"fast",if_else(growth >10,"regular",if_else(growth > 5,"slow","very slow"))),"effeciency"=harvest/growth)

i<-ggplot(data2,aes(size, group=g_class))
i+geom_boxplot(aes(fill=g_class))+coord_flip()+labs(x="Size",title = "Boxplots of Size by Growth Group")+scale_fill_discrete(name="Growth Group")
```
Observing our parrallel boxplots we see that the average growth speed has the largest size with the very slow speed have a much smaller size. We also see that the fast speed have a small size similar. It may be interesting to see if there is a significant difference statistically between any of the groups.



Then we can create a barplot with information on firmness counts subdivided by growth class
```{r bar1}
j<- ggplot(data2,aes(firmness))
j+geom_bar(aes(fill=g_class),position = "dodge")+labs(x="Firmness",y="Quantity",title="Barplot of Firmness grouped by Growth Group")+scale_fill_discrete(name="Growth Group")
```
The bar graph does not seem to reveal any replationship between the firmness of the berries and the growth group to which the berry belongs.


Next we can look at some contingency tables of our data, to make things interesting we can create 2 way and even 3 way contingency tables. Then we can look at some numerical sumarise of the numeric variables when grouped by each of the categorical variables individually.
```{r tables}
data2 %>% select(natural_gift,firmness) %>% table()
data2 %>% select(natural_gift,firmness,g_class) %>% table()

data2 %>% group_by(firmness) %>% summarise(avg_growth=mean(growth),avg_size=mean(size),avg_harvest=mean(harvest))
data2 %>% group_by(g_class) %>% summarise(avg_size=mean(size),avg_harvest=mean(harvest))#g_class based upon growth
data2 %>% group_by(natural_gift) %>% summarise(avg_growth=mean(growth),avg_size=mean(size),avg_harvest=mean(harvest))
```
The contingency tables don't show very much likely because the number of berries is small, but future games may include more berries and improve the results. The summary statistics for each of the 3 categorical variables seem interesting. It may be worth while to see if a growth class(or Growth) or firmness can be predicted using some of the numeric variables.

#Pokemon
Finally we may want to look at some data on the pokemon like the type, the height, and even the weight.
```{r pokemonData}
pokemons <- API_Wrapper("pokemon",type="pokemon",limit=10000)[,1]

type <- lapply(pokemons, API_Wrapper,fun="pokemon",type="pokemon",val="types",val2="type",val3="name")
type <- lapply(type, `[`,1)
type <- depth2(type)

height <- lapply(pokemons, API_Wrapper,fun="pokemon",type="pokemon",val="height")
height <- ResultsToDataframe(height)

weight <- lapply(pokemons, API_Wrapper,fun="pokemon",type="pokemon",val="weight")
weight <- ResultsToDataframe(weight)

data3 <- cbind(type,height,weight)
colnames(data3)<- c("type","height","weight")
```

We can even look at a barplot of the various types of pokemon.
```{r bar2}
k<-ggplot(data3,aes(type))
k+geom_bar(aes(fill=type))+theme(axis.text.x = element_text(angle=45))+labs(x="Pokemon Main Type",y="Count",title="Quantity in each Pokemon Main Type")+guides(fill="none")
```
observing the barplot we see that water and normal are the developers favorite main pokemon types. This may or may not be due to pokemon in ocean locations have exclusively water main elements. It may be interesting to see if this is the reason for the large number of water type pokemon.


Finally we can create a scatterplot of height vs weight grouped by the pokemon type, we specifically exclude the pokemon with height values over 200 to get a better look at the data
```{r scatter2}
data4 <- data3 %>% filter(height < 200)
l <- ggplot(data4,aes(height,weight))
l+geom_point(aes(color=type))+labs(title="Height vs Weight by Main Type",x="Height",y="Weight")+theme(legend.key.height = unit(.35, 'cm'))
```
Here we see the data is somewhat difficult to interpret so it may be beneficial to look at each type individually. It also may be interesting to see if a prediction can be made for height using type and weight.


